Calculating Module Enrichment and Visualizing Data on Large-scale Molecular Maps with the R packages \pkg{ACSNMineR} and \pkg{RNaviCell}

### 2016-01-10 jss/article.tex
[skip ci] update examples.

\begin{verbatim}
# analysis mtsig signature file

@@

\begin{verbatim}
apply(eset[probe_list,pdat$er == 0],1,mean) -> er_minus_mean
names(er_minus_mean) <- symbol_list
er_minus_mean <- as.matrix(er_minus_mean)
colnames(er_minus_mean) <- c('exp')

@@

>>>>

\begin{verbatim}
# Import MSigDB canonical pathways and calculate enrichment on this database

@@

At last, we visualize the mean expression values for ER negative samples for all
genes differentially expressed on the ACSN master map using \code{ACSNMineR}
commands to create heatmaps.

\begin{verbatim}
# Select ER negative samples and calculate mean expression values

apply(eset[probe_list,pdat$er == 0],1,mean) -> er_minus_mean
names(er_minus_mean) <- symbol_list
er_minus_mean <- as.matrix(er_minus_mean)
colnames(er_minus_mean) <- c('exp')

# create a NaviCell session, import the expression matrix on the map and create
# heatmaps to represent the data points.

@@

The Figure \ref{fig:mainz} is displaying the map for gene that having a
corrected p-value $< 0.05$ and a log fold-change $>1$. we can see that the genes
are concentrated in the regions of the map corresponding to the cell cycle, cell
motility, apoptosis and survival.

\begin{figure}[!ht]
  \caption{Mean expression values for ER negative differentially expressed
  genes in breast cancer visualized as heatmaps on the ACSN master map.} 
  \centering
  \includegraphics[width=0.8\textwidth]{figures/mainz_acsn.pdf}
  \label{fig:mainz}
\end{figure}

\subsection{Analysis of glioblastoma mutation frequencies}

\begin{table}[h!]
  \centering
  \caption{MSigDB canonical pathway database enrichment for genes differentially expressed between ER
positive and ER negative samples in breast cancer.  Module : name of the
module. Mod. size: size of the module. Nb genes: number of genes from input
which are found in the module. Cor. pval: corrected p-value.}
\label{tab:table_gbm}

\begin{tabular}{l|c|c|c}
\hline
module & Mod. size & Nb genes & Cor. pval\\ 
\hline
CellCycle:G2\_M\_CHECKPOINT & 14 & 8 & 3.92e-02\\
CellCycle:S\_PHASE\_CHECKPOINT & 6 & 5 & 3.92e-02\\
DNA\_repair:S\_PHASE\_CHECKPOINT & 45 & 19 & 1.58e-02\\
EMT\_motility & 621 & 176 & 1.06e-03\\
EMT\_motility:CELL\_MATRIX\_ADHESIONS & 72 & 44 & 2.55e-11\\
EMT\_motility:ECM & 137 & 64 & 2.34e-09\\
EMT\_motility:EMT\_REGULATORS & 615 & 173 & 1.49e-03\\
EMT\_motility:CYTOSKELETON\_POLARITY & 154 & 47 & 3.92e-02\\
EMT\_motility:DESMOSOMES & 29 & 15 & 4.62e-03\\
ACSN\_master:CELL\_MATRIX\_ADHESIONS & 72 & 44 & 2.55e-11\\
ACSN\_master:DESMOSOMES & 29 & 15 & 4.62e-03\\
ACSN\_master:S\_PHASE\_CHECKPOINT & 46 & 20 & 8.23e-03\\
ACSN\_master:ECM & 138 & 64 & 2.56e-09\\
ACSN\_master:EMT\_REGULATORS & 625 & 177 & 1.06e-03\\
ACSN\_master:CYTOSKELETON\_POLARITY & 154 & 47 & 3.92e-02\\
Survival:ECM & 23 & 11 & 3.92e-02\\
\hline
\end{tabular}
\end{table}

\begin{figure}[!ht]
  \caption{Mean expression values for ER negative differentially expressed
  genes in breast cancer visualized as heatmaps on the ACSN master map.} 
  \centering
  \includegraphics[width=0.8\textwidth]{figures/gbm.pdf}
  \label{fig:mainz}
\end{figure}
